The Negative Effect of Traffic Noise on House Prices: A Landscape Hedonic Analysis

### 2013-10-15 reports/Draft.Rnw
final edits. submitted DraftAdobe and TitlePageAdobe to the Journal of Regional Science

\documentclass{article}
\usepackage[round]{natbib}
\usepackage{caption, amsmath, graphicx, setspace, multirow, color, hyperref, array}

@@

\title{The Negative Effect of Traffic Noise on House Prices: A Landscape Hedonic Analysis}
\date{\today}

\doublespacing
\begin{document}
\maketitle
\begin{singlespace}
\begin{abstract}
One consequence of the expanding road network and its associated traffic is increased levels of traffic noise.  While the hedonic literature has consistently shown a negative effect of this phenomenon on the real estate market, research in the United States has often relied on crude measures of traffic noise. Here, we reduce the measurement error of traffic noise exposure through a detailed model of noise propagation over the landscape. Additionally, we estimate the impact on single family home transactions throughout the St.\ Paul, Minnesota, urban area using spatially explicit local regression techniques to allow for spatial non-stationarity in the hedonic function. Contrary to previous work, we find no evidence that the effect of traffic noise varies significantly across the several hundred square miles of our study region, nor by traffic noise levels. We do find significant differences in the impact of traffic noise before and after the economic recession of 2008-09. Our results suggest that an increase in traffic noise of one decibel decreases house prices by an average of 0.19 percent before September 2008 and an average of 0.37 percent after September 2008.
\end{abstract}
\end{singlespace}

\newpage
\section{Motivation and Past Research}\label{sec:lit}
Prolonged exposure to traffic noise affects people in a number of ways, ranging from simple annoyance \citep{Miedema2001, Ouis2001, Ohrstrom2007, DeKluizenaar2013, Weinhold2013}, to sleep disturbance \citepalias{Netherlands2004}, to increasing risk for stroke \citep{Sorensen2011}, hypertension \citep{Jarup2008, Bodin2009}, myocardial infarction \citep{Babisch2005}, and overall quality of life \citep{Shepherd2013}. The noise level at which such effects are observed does not have to be high.  It has been shown that people exposed to traffic noise with a 24-hour average of 55 decibels (dBA) are found to be at a higher risk for hypertension \citep{Barregard2009, Bodin2009}, and those exposed to 60 dBA or greater are found to be at a higher risk for stroke \citep{Sorensen2011}.

@@

\begin{table}[htb]
\caption{Variable Description and Summary Statistics}\label{tab:sumStats}
\makebox[\linewidth][c]{
\small
\begin{tabular}{lrrrrrrr}

@@

\subsubsection{Does the impact of traffic noise vary over space?}
In the previous simulation we found that the hedonic regression function indeed exhibits spatial non-stationarity. However, we have not yet established which variables have spatially varying coefficients. In the second simulation, we seek to explore which of the hedonic regression coefficients exhibit spatial non-stationarity. We repeat the Monte Carlo simulation described earlier, but only at the optimal bandwidth from the original LWR model. We collected the standard deviation of the estimated LWR coefficients for each variable in each iteration of the simulation. The intuition of this simulation is as follows: a variable might truly exert a spatially \emph{stationary} effect on house prices, but our estimates vary over space due to chance. If this is the case, by shuffling the location of our observations and rerunning the LWR model, we should see a similar dispersion in regression coefficients over space with the reshuffled data as with our actual data. However, if the coefficient exhibits true spatial \emph{non-stationarity}, it is unlikely that we will see such a large of a standard deviation in the coefficient estimates after having reshuffled the data (because now the houses for which a coefficient exerts a strong effect are probably no longer near each other).

@@

In this section we look for evidence of temporal non-stationarity in the hedonic coefficient on traffic noise within our data. Specifically, we create two new variables, ``\texttt{months since September 2008}'' and a dummy variable ``\texttt{post}=1'' for all sales after September 2008. We then regress our LWR noise coefficients on these two variables and their interaction term as shown in equation \eqref{eq:LWRinteraction}, 
\begin{equation}\label{eq:LWRinteraction}
\hat{\beta}_{LWR} = \alpha _0 + \alpha _1 \texttt{*Months since Sep. 2008} + \alpha _2  \texttt{*Post} + \alpha _3 \texttt{*Post*Months since Sep. 2008} + \epsilon.
\end{equation}
The results of this regression will help us understand the patterns of change in the implicit price of traffic noise. In particular, $\alpha_1 \neq 0$ suggests a linear temporal trend in the marginal willingness to pay to avoid traffic noise, while $\alpha_2 \neq 0$ suggests that there was a structural break after September 2008, and $\alpha_3 \neq 0$ will imply a change in the temporal trend after September 2008. The linear regression results presented in Table \ref{tab:betaMAX} show that there the estimated LWR coefficients for traffic noise are significantly smaller after September 2008.
\begin{table}[htb]
\caption{Dependent Variable = Traffic Noise LWR Coefficients}\label{tab:betaMAX}
\begin{center}
\begin{tabular}{lrrr}
           & Estimate & Std.\ Err. & t-stat  \\  \hline
(Intercept) & -1.7e-03 &  4.1e-05 &  -43.4   \\
Months since Sep 2008       & 6.9e-06 &  2.0e-06  &   3.5    \\
Post        &-8.0e-04 &  7.2e-05  & -11.1    \\
Post*Months since Sep 2008  &-8.9e-05 &  4.5e-06  & -19.7    \\ \hline
\multicolumn{4}{l}{Residual standard error: 0.003} \\
\multicolumn{4}{l}{N = 31,748, R$^2$: 0.093} \\ %,  Adjusted R$^2$: 0.093
%\multicolumn{4}{l}{Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1}
\end{tabular}
\end{center}
\end{table}

@@

\begin{table}[ht]
\begin{center}
\caption{Noise Coefficient Estimates in Dakota County with and without Additional Structural Variables}\label{tab:Dak}
\begin{tabular}{lll}
 & \multicolumn{2}{p{1.25in}}{Additional Variables Included in LWR?} \\
Locally Weighted Regression Model & \multicolumn{1}{c}{no} & \multicolumn{1}{c}{yes} \\ \cline{1-3}
\multirow{2}{*}{Model (1) = Structural Variables} & -0.0015 & -0.0015 \\ 
   & (0.0023) & (0.0031)  \\[.2cm]
\multirow{2}{*}{Model (2) = Model (1) + Locational Variables} & -0.0014 & -0.0014  \\ 
   & (0.0018) & (0.0022)  \\[.2cm] 
\multirow{2}{*}{Model (3) = Model (2) + City Fixed Effects} & -0.0014 & -0.0014 \\ 
   & (0.0022) & (0.0018) \\[.2cm] \cline{1-3}
\multicolumn{3}{p{4.25in}}{Mean and (standard deviation) LWR traffic noise coefficients when the number of bedrooms, bathrooms, and garage size are included as additional control variables in the regression.}
\end{tabular}
\end{center}
\end{table}

@@

\begin{singlespace}
\bibliographystyle{plainnat}
\bibliography{NoiseBibliography}
\end{singlespace}
\end{document}

>>>>

\documentclass{article}
\usepackage[round]{natbib}
\usepackage[nolists]{endfloat}
\usepackage[width = 5in]{geometry}
\usepackage{caption, amsmath, graphicx, setspace, multirow, color, hyperref, array}

%\renewcommand{\efloatseparator}{\mbox{}}

@@

\title{The Negative Effect of Traffic Noise on House Prices: A Landscape Hedonic Analysis}
\date{}

\doublespacing
\begin{document}
\maketitle
\pagenumbering{roman}
\begin{singlespace}
\begin{abstract}
One consequence of the expanding road network and its associated traffic is increased levels of traffic noise.  While the hedonic literature has consistently shown a negative effect of this phenomenon on the real estate market, research in the United States has often relied on crude measures of traffic noise. Here, we reduce the measurement error of traffic noise exposure through a detailed model of noise propagation over the landscape. Additionally, we estimate the impact on single family home transactions throughout the St.\ Paul, Minnesota, urban area using spatially explicit local regression techniques to allow for spatial non-stationarity in the hedonic function. Contrary to previous work, we find no evidence that the effect of traffic noise varies significantly across the several hundred square miles of our study region, nor by traffic noise levels. We do find significant differences in the impact of traffic noise before and after the economic recession of 2008-09. Our results suggest that an increase in traffic noise of one decibel decreases house prices by an average of 0.19 percent before September 2008 and an average of 0.37 percent after September 2008.
\end{abstract}
\end{singlespace}

\clearpage
\pagenumbering{arabic}

\section{Motivation and Past Research}\label{sec:lit}
Prolonged exposure to traffic noise affects people in a number of ways, ranging from simple annoyance \citep{Miedema2001, Ouis2001, Ohrstrom2007, DeKluizenaar2013, Weinhold2013}, to sleep disturbance \citepalias{Netherlands2004}, to increasing risk for stroke \citep{Sorensen2011}, hypertension \citep{Jarup2008, Bodin2009}, myocardial infarction \citep{Babisch2005}, and overall quality of life \citep{Shepherd2013}. The noise level at which such effects are observed does not have to be high.  It has been shown that people exposed to traffic noise with a 24-hour average of 55 decibels (dBA) are found to be at a higher risk for hypertension \citep{Barregard2009, Bodin2009}, and those exposed to 60 dBA or greater are found to be at a higher risk for stroke \citep{Sorensen2011}.

@@

\begin{table}
\caption{Variable Description and Summary Statistics}\label{tab:sumStats}
\makebox[\linewidth][c]{
\small
\begin{tabular}{lrrrrrrr}

@@

\subsection{Does the Implicit Price of Traffic Noise Change over Space?}
In the previous simulation we found that the hedonic regression function indeed exhibits spatial non-stationarity. However, we have not yet established which variables have spatially varying coefficients. In the second simulation, we seek to explore which of the hedonic regression coefficients exhibit spatial non-stationarity. We repeat the Monte Carlo simulation described earlier, but only at the optimal bandwidth from the original LWR model. We collected the standard deviation of the estimated LWR coefficients for each variable in each iteration of the simulation. The intuition of this simulation is as follows: a variable might truly exert a spatially \emph{stationary} effect on house prices, but our estimates vary over space due to chance. If this is the case, by shuffling the location of our observations and rerunning the LWR model, we should see a similar dispersion in regression coefficients over space with the reshuffled data as with our actual data. However, if the coefficient exhibits true spatial \emph{non-stationarity}, it is unlikely that we will see such a large of a standard deviation in the coefficient estimates after having reshuffled the data (because now the houses for which a coefficient exerts a strong effect are probably no longer near each other).

@@

In this section we look for evidence of temporal non-stationarity in the hedonic coefficient on traffic noise within our data. Specifically, we create two new variables, ``\texttt{months since September 2008}'' and a dummy variable ``\texttt{post}=1'' for all sales after September 2008. We then regress our LWR noise coefficients on these two variables and their interaction term as shown in equation \eqref{eq:LWRinteraction}, 
\begin{equation}\label{eq:LWRinteraction}
\hat{\beta}_{LWR} = \alpha _0 + \alpha _1 \texttt{*Months since Sep. 2008} + \alpha _2  \texttt{*Post} + \alpha _3 \texttt{*Post*Months since Sep. 2008} + \epsilon.
\end{equation}
The results of this regression will help us understand the patterns of change in the implicit price of traffic noise. In particular, $\alpha_1 \neq 0$ suggests a linear temporal trend in the marginal willingness to pay to avoid traffic noise, while $\alpha_2 \neq 0$ suggests that there was a structural break after September 2008, and $\alpha_3 \neq 0$ will imply a change in the temporal trend after September 2008. The linear regression results presented in Table \ref{tab:betaMAX} show that there the estimated LWR coefficients for traffic noise are significantly smaller after September 2008.
\begin{table}
\caption{Dependent Variable = Traffic Noise LWR Coefficients}\label{tab:betaMAX}
\begin{center}
\begin{tabular}{lrrr}
           & Estimate & Std.\ Err. & t-stat  \\  \hline
(Intercept) & -1.7e-03 &  4.1e-05 &  -43.4   \\
Months since Sep 2008       & 6.9e-06 &  2.0e-06  &   3.5    \\
Post        &-8.0e-04 &  7.2e-05  & -11.1    \\
Post*Months since Sep 2008  &-8.9e-05 &  4.5e-06  & -19.7    \\ \hline
\multicolumn{4}{l}{Residual standard error: 0.003} \\
\multicolumn{4}{l}{N = 31,748, R$^2$: 0.093} \\ %,  Adjusted R$^2$: 0.093
%\multicolumn{4}{l}{Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1}
\end{tabular}
\end{center}
\end{table}

@@

\begin{table}
\begin{center}
\caption{Noise Coefficient Estimates in Dakota County with and without Additional Structural Variables}\label{tab:Dak}
\begin{tabular}{lll}
 & \multicolumn{2}{p{1.25in}}{Additional Variables Included in LWR?} \\
Locally Weighted Regression Model & \multicolumn{1}{c}{no} & \multicolumn{1}{c}{yes} \\ \cline{1-3}
\multirow{2}{*}{Model (1) = Structural Variables} & -0.0015 & -0.0015 \\ 
   & (0.0023) & (0.0031)  \\[.2cm]
\multirow{2}{*}{Model (2) = Model (1) + Locational Variables} & -0.0014 & -0.0014  \\ 
   & (0.0018) & (0.0022)  \\[.2cm] 
\multirow{2}{*}{Model (3) = Model (2) + City Fixed Effects} & -0.0014 & -0.0014 \\ 
   & (0.0022) & (0.0018) \\[.2cm] \cline{1-3}
\multicolumn{3}{p{4.25in}}{Mean and (standard deviation) LWR traffic noise coefficients when the number of bedrooms, bathrooms, and garage size are included as additional control variables in the regression.}
\end{tabular}
\end{center}
\end{table}

@@

\newpage
\begin{singlespace}
\bibliographystyle{plainnat}
\bibliography{NoiseBibliography}
\end{singlespace}
\end{document}
